Experiments on ANTS Algorithms Without Communication

### 2014-12-09 paper/draft.tex
revised sections 3-5

Given the number of agents taking part in the search, we could imagine an algorithm in which the grid in divided into $k$ parts, similar to a pizza being divided into $k$ equal parts. Each agent could cover his portion, and this would yield an asymptotically optimal algorithm. However, this assumes that the agents can communicate with each other and decide which portion belongs to which agent.

@@

This is what the spiral algorithms does when its non-uniform in $k$.

@@

Once we have the spiral algorithm non-uniform in $k$, we can make it uniform in $k$ by successively doubling estimates of $k$. This means that at each stage, we take a guess of what $k$ is, and we simulate the algorithm. There is some additional bookkeeping to ensure that enough iterations are made, but the optimization is similar. There is higher lower bound on how fast these algorithms can run, and we get a multiplicative factor of $f(\log k)$, where $f$ is a function such that $\sum \dfrac{1}{f(n)}$ converges. There is a corresponding lower bound that shows this algorithm is asymptotically optimal.

@@

The lines algorithm is the study of \cite{lenzen2014trade}, where they cover a lines algorithm which is non-uniform in $D$ and non-uniform in $k$. We use the ideas of \cite{feinerman2012collaborative} to design a lines algorithm which is uniform in both $k$ and $D$. The lines algorithm has many optimizations since \cite{lenzen2014trade} are trying to reduce \emph{selection complexity}. Selection complexity is a metric which measures how likely an algorithm is to occur naturally in nature. Of course, this notion is very vague, but formally, the selection complexity $\chi$ of an algorithm $A$ is $\chi(A) = b + \log l$ where $b$ is the amount of memory usage and $l$ is smallest $l$ for which the algorithm does not need finer probabilities than $2^{-l}$. In \cite{lenzen2014trade}, Lenzen et al. provided a lines algorithm with selection complexity $O(\log \log D)$.

@@

\begin{codebox}
\Procname{Each agent performs the following iteration}
\li dir[0] $\leftarrow$ up, dir[1] $\leftarrow$ down
\li $i \leftarrow C_{\frac{1}{2}}$
\li moveDirection($i$)
\li \If $C_{\frac{1}{2}} = 0$ \Then
\li moveDirection(left)
\li \Else moveDirection(right) \End \End
\li return to origin
\end{codebox}

@@

% TO DO
The sketch of the proof of correctness is the following ...????

@@

The main idea for achieving uniformity in $k$ is borrowed from \cite{feinerman2012collaborative}: we start with a small guess for $k$, run the algorithm uniform in $D$ under this assumption, and then double out guess for $k$ and repeat. The implementation actually interleaves the increments of the $D$ and $k$ estimates, running a certain number of iterations of the non-uniform subroutine for each estimate.

>>>>

Given the number of agents taking part in the search, we could imagine an algorithm in which the grid in divided into $k$ parts, similar to a pizza being divided into $k$ equal parts. Each agent could cover its portion, and this would yield an asymptotically optimal algorithm. However, this assumes that the agents can communicate with each other and decide which portion belongs to which agent.

@@

This is what the spiral algorithms does when it's non-uniform in $k$.

@@

Once we have the spiral algorithm non-uniform in $k$, we can make it uniform in $k$ by successively doubling estimates of $k$. This means that at each stage, we take a guess of what $k$ is, and we simulate the algorithm given the guess. There is some additional bookkeeping to ensure that enough iterations are made, but the optimization is similar. There is higher lower bound on how fast these algorithms can run, and we get a multiplicative factor of $f(\log k)$, where $f$ is a function such that $\sum \dfrac{1}{f(n)}$ converges. There is a corresponding lower bound that shows this algorithm is asymptotically optimal.

@@

The lines algorithm is the study of \cite{lenzen2014trade}, where they cover a lines algorithm that is non-uniform in $D$ and one that is uniform in $D$ but non-uniform in $k$. We use the ideas of \cite{feinerman2012collaborative} to design a lines algorithm which is uniform in both $k$ and $D$. The lines algorithm has many optimizations since \cite{lenzen2014trade} are trying to reduce \emph{selection complexity}. Selection complexity is a metric which measures how likely an algorithm is to occur in nature. Of course, this notion is very vague, but formally, the selection complexity $\chi$ of an algorithm $A$ is $\chi(A) = b + \log l$ where $b$ is the amount of memory usage and $l$ is smallest $l$ for which the algorithm does not need finer probabilities than $2^{-l}$. In \cite{lenzen2014trade}, Lenzen et al. provided a lines algorithm with selection complexity $O(\log \log D)$, as opposed to $O(\log D)$ for the spiral algorithms and $\infty$ for harmonic search.

@@

\begin{codebox}
\Procname{Each agent performs the following iteration}
\li \If $C_{\frac{1}{2}} = 0$ \Then
\li moveDirection(up)
\li \Else moveDirection(down) \End
\li \If $C_{\frac{1}{2}} = 0$ \Then
\li moveDirection(left)
\li \Else moveDirection(right) \End \End
\li return to origin
\end{codebox}

@@

@@

The main idea for achieving uniformity in $k$ is borrowed from \cite{feinerman2012collaborative}: we start with a small guess for $k$, run the algorithm uniform in $D$ under this assumption, and then double our guess for $k$ and repeat. The implementation actually interleaves the increments of the $D$ and $k$ estimates, probabilistically running a number of iterations of the non-uniform subroutine for each estimate based on the outcomes of a series of coin flips, as explained in the subsection above.

The expected time complexity of this algorithm is the same as that for the uniform variant of the spirals algorithms.
